---
title: "SongmeterAnalysis"
author: "Emily"
date: "`r format(Sys.time(), '%d %B, %Y')`"
summary: "This is analysis of acoustic environment at nests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(pacman)
p_load(dplyr,data.table,multcomp,ggplot2,ggrepel,MuMIn)
```

###Run SongmeterDataInput.R first


```{r include=FALSE}

#Set directory:
workingdir <- c("/Users/ehudson/Google Drive/My Drive/fka_Dropbox/SongmeterProj/")


wcsppbs=read.csv(paste0(workingdir,"wcsppbs.csv"),na.string="#N/A")
gcsppbs=read.csv(paste0(workingdir,"gcsppbs.csv"),na.string="#N/A")
allnestprops=read.csv(paste0(workingdir,"WCSPproportions1.csv"),na.string="#N/A")
```
```{r include=FALSE}
wcsppbs_all=read.csv(paste0(workingdir,"wcsppbs_all.csv"))
gcsppbs_all=read.csv(paste0(workingdir,"gcsppbs_all.csv"))
nesttotals=read.csv(paste0(workingdir,"TotalSongsPerNest.csv"))
nestrespdiff=read.csv(paste0(workingdir,"NestRespDiffs.csv"))
```
```{r include=FALSE}
nestcols = rainbow(length(unique(wcsppbs$Nest.number)))
nestnum = length(unique(wcsppbs$Nest.number))
```
```{r include=FALSE}
## Add nestling sex
nestsex = read.csv(paste0(workingdir,"HAPANestlPlayback_complete2.csv"),colClasses = "character", na.string="#N/A")

nestsex = nestsex[nestsex$use != "no",]
nestsex = nestsex[nestsex$Species != "Other",]
nestsex = nestsex[nestsex$Sex != "",]
nestsex = nestsex[nestsex$Year == "2017",]
#nestsex$Sex <- na.omit(nestsex)
gcsppbs$Sex <- nestsex$Sex[match(gcsppbs$Chick, nestsex$Chick)]
gcsppbs <- gcsppbs[!is.na(gcsppbs$Sex),]
gcsppbs$Sex <- as.factor(gcsppbs$Sex)
gcsppbs$WCrate <-nesttotals$WCrate[match(gcsppbs$Nest.number,nesttotals$Nest)]
gcsppbs$GCrate <-nesttotals$GCrate[match(gcsppbs$Nest.number,nesttotals$Nest)]

wcsppbs$Sex <- nestsex$Sex[match(wcsppbs$Chick, nestsex$Chick)]
wcsppbs <- wcsppbs[!is.na(wcsppbs$Sex),]
```
```{r include=FALSE}
#This filters out 1746:
#wcsppbs <- wcsppbs[!is.na(wcsppbs$Sex),]

wcsppbs$Sex <- as.factor(wcsppbs$Sex)
wcsppbs$WCrate <- nesttotals$WCrate[match(wcsppbs$Nest.number,nesttotals$Nest)]
wcsppbs$GCrate <- nesttotals$GCrate[match(wcsppbs$Nest.number,nesttotals$Nest)]

wcsppbs$condition <- as.numeric(as.character(wcsppbs$mass))/as.numeric(as.character(wcsppbs$tarsus))
gcsppbs$condition <- as.numeric(as.character(gcsppbs$mass))/as.numeric(as.character(gcsppbs$tarsus))


wcsppbs$ChirpGC <- gcsppbs$Track[match(wcsppbs$Nest.number, gcsppbs$Nest.number)]
wcsppbs$Chirpdiff <- wcsppbs$ChirpGC-wcsppbs$Track
wcsppbs$Chirpprop <-  wcsppbs$Track/wcsppbs$ChirpGC

wcsppbs$Chirpprop <- replace(wcsppbs$Chirpprop,wcsppbs$Chirpprop=="Inf","0") #When no chirps to WC, chirpprop is zero

#And without excluding nonresponders:
gcsppbs_all$Sex <- nestsex$Sex[match(gcsppbs_all$Chick, nestsex$Chick)]
gcsppbs_all <- gcsppbs_all[!is.na(gcsppbs_all$Sex),]
gcsppbs_all$Sex <- as.factor(gcsppbs_all$Sex)
wcsppbs_all$Sex <- nestsex$Sex[match(wcsppbs_all$Chick, nestsex$Chick)]
wcsppbs_all <- wcsppbs_all[!is.na(wcsppbs_all$Sex),]
wcsppbs_all$Sex <- as.factor(wcsppbs_all$Sex)
```
```{r include=FALSE}
nestmeans<-aggregate(Track~Nest.number,mean,data=wcsppbs)

nestmeans$Nest.number<-factor(nestmeans$Nest.number)
gcsppbs1 <- as.data.frame(gcsppbs[,c("band.number","Nest.number","Clutch","Pre","Track","Feather","WCSPprop","Sex","WCrate","GCrate","condition","mass")])
gcsppbs1$pb <- "GCSP"
wcsppbs1<-as.data.frame(wcsppbs[,c("band.number","Nest.number","Clutch","Pre","Track","Feather","WCSPprop","Sex","WCrate","GCrate","condition","mass")])
wcsppbs1$Feather <- as.numeric(as.character(wcsppbs1$Feather))
wcsppbs1$pb <- "WCSP"
allpbs <- rbind(wcsppbs1,gcsppbs1) #feather messed up
allpbsM <-droplevels(subset(allpbs,allpbs$Sex=="M"))
allpbsF <-droplevels(subset(allpbs,allpbs$Sex=="F"))
nestprops <- as.data.frame(nestrespdiff)
nestprops$Exp <- allnestprops$WCSPprop[match(nestprops$Nest,allnestprops$Nest)]
nestprops$WCRespAvg <- nestmeans$Track[match(nestprops$Nest,nestmeans$Nest.number)]

```
```{r include=FALSE}
nestmeans$WCSPprop<-aggregate(WCSPprop~Nest.number,max,data=wcsppbs)$WCSPprop
nestmeans$Clutch <- aggregate(Clutch~Nest.number,mean,data=wcsppbs)$Clutch

##With proportion of responses to WCSP/total as response

nestprops <- as.data.frame(nestrespdiff)
nestprops$Exp <- allnestprops$WCSPprop[match(nestprops$Nest,allnestprops$Nest)]
summary(lm(nestprops$RespProp~nestprops$Exp))
plot(nestprops$RespProp~nestprops$Exp)

summary(lm(nestprops$RespDiff~nestprops$Exp))
plot(nestprops$RespDiff~nestprops$Exp)
#nestprops$WCRespAvg <- nestmeans$Track[match(nestprops$Nest,nestmeans$Nest.number)] 

nestmeans_all<-aggregate(Track~Nest.number,mean,data=wcsppbs_all)

nestmeans_all$Nest.number<-factor(nestmeans_all$Nest.number)

nestmeans_all$WCSPprop<-aggregate(WCSPprop~Nest.number,max,data=wcsppbs_all)$WCSPprop
nestmeans_all$Clutch <- aggregate(Clutch~Nest.number,mean,data=wcsppbs_all)$Clutch
```
```{r include=FALSE}

#nestmeans<-aggregate(Response~Nest.number,mean,data=wcsppbs)
nestmeans$Nest.number<-factor(nestmeans$Nest.number)

nestmeans$WCSPprop<-aggregate(WCSPprop~Nest.number,max,data=wcsppbs)$WCSPprop

#add condition column 
wcsppbs$condition <- as.numeric(wcsppbs$mass)/as.numeric(wcsppbs$tarsus)
gcsppbs$condition <- as.numeric(gcsppbs$mass)/as.numeric(gcsppbs$tarsus)

```



```{r echo=FALSE}
# Make separate male and female dataframe
gcsppbsM <- subset(gcsppbs, gcsppbs$Sex=="M")
gcsppbsF <- subset(gcsppbs, gcsppbs$Sex=="F")
wcsppbsM <- subset(wcsppbs, wcsppbs$Sex=="M")
wcsppbsF <- subset(wcsppbs, wcsppbs$Sex=="F")
nestmeans_no40 <- nestmeans_all[-c(17),]
nestmeans <- nestmeans_no40
```
```{r echo=FALSE}
ggplot(data=nestmeans_all,aes(x=jitter(WCSPprop),y=Track,col=Nest.number,label=Nest.number))+geom_point(aes(fill=Nest.number))+geom_text_repel()+labs(title="without excluding nonresponders")
write.csv(nestmeans,"NestMeans.csv")
write.csv(allpbs,"All2017PBdata.csv")
ggsave("New_nests_all.png")


```
```{r echo=FALSE}
g1<-ggplot(data=nestmeans_all,aes(x=jitter(WCSPprop),y=Track))+geom_point(fill="grey",shape=21,size=3,stroke=1)+labs(x="Proportion of white-crowned sparrow song \n relative to total Zonotrichia song ", y="Average nest-wide response to  \n white-crowned sparrow playback",size=10)#without excluding nonresponders
g1+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA),
    axis.line = element_line(colour = "black"),
  axis.text=element_text(size=12)
)
anova(lm(nestmeans_all$WCSPprop~nestmeans_all$Track))
ggsave("New_nests_all_grey.png")


```
```{r echo=FALSE}
g1<-ggplot(data=nestmeans,aes(x=jitter(WCSPprop),y=Track))+geom_point(fill="grey",shape=21,size=3,stroke=1)+labs(x="Proportion of white-crowned sparrow song \n relative to total Zonotrichia song ", y="Average nest-wide response to  \n white-crowned sparrow playback",size=10) #excluding nest 40
g1+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA),
    axis.line = element_line(colour = "black"),
  axis.text=element_text(size=12)
)
anova(lm(nestmeans$WCSPprop~nestmeans$Track))
ggsave("New_nests_no40_grey.png")
```
```{r echo=FALSE}
nestmeans_few <- subset(nestmeans_all,nestmeans_all$Nest.number != 1740)
nestmeans_few <- droplevels(subset(nestmeans_few,nestmeans_few$Nest.number != 1733))
write.csv(nestmeans_few,"FinalNestMeans.csv")
summary(lm(nestmeans_few$Track~nestmeans_few$WCSPprop))

#table(nestmeans_few$Nest.number)

summary(lm(nestmeans$Track~nestmeans$WCSPprop))
```
```{r echo=FALSE}
summary(lm(nestmeans$Track~nestmeans$WCSPprop))

```

##1. Analyzing all data, with playback type, pre-track response, sex and feather length as covariates. 



```{r echo=FALSE}
library(MuMIn)



print("No interaction, then with playback*WCproportion and playback*Sex interactions.")
lm1 <- lm(Track~pb+WCSPprop+Pre+Sex+as.numeric(Feather),data=allpbs)
lmred<- lm(Track~pb+WCSPprop+Pre+as.numeric(Feather),data=allpbs)
anova(lm1,lmred)
summary(lm1)
lm2 <- lm(Track~pb*WCSPprop+Pre+Sex+as.numeric(Feather),data=allpbs)
lm2red <- lm(Track~Pre+Sex+as.numeric(Feather),data=allpbs)
anova(lm2,lm2red)
summary(lm2)

lm3 <- lm(Track~pb*Sex+Pre+WCSPprop+as.numeric(Feather),data=allpbs)
lm3red <- lm(Track~Pre+WCSPprop+as.numeric(Feather),data=allpbs)
anova(lm3,lm3red)
summary(lm3)


allpbs <- subset(allpbs,allpbs$Sex!="NA")
allpbs$figtrt <- as.factor(paste(allpbs$Sex,allpbs$pb,sep = ""))
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

Including clutch size; WC proportion now drops to p=0.051.

```{r echo=FALSE}
lm4 <- lm(Track~pb+Clutch+Sex+Pre+WCSPprop+as.numeric(Feather),data=allpbs)
summary(lm4)
allpbs <- subset(allpbs,allpbs$Sex!="NA")
allpbs$figtrt <- as.factor(paste(allpbs$Sex,allpbs$pb,sep = ""))
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

levels(allpbs$figtrt) <- c("Female Nestlings GC playback","Female Nestlings WC playback","Male Nestlings GC playback","Male Nestlings WC playback")

fig1d <- ggplot(data=allpbs,aes(y=Track,x=WCSPprop,shape=figtrt,fill=figtrt),colour=figtrt)+geom_point()+geom_smooth(method=lm, se=FALSE, fullrange=TRUE,aes(col=figtrt))+scale_shape_manual(values=c(21:24))+scale_color_manual(values=gg_color_hue(4))
fig1d=fig1d+theme(
  panel.background = element_rect(fill = "transparent",colour = NA), # or element_blank
  #panel.grid.minor = element_blank(), 
  #panel.grid.major = element_blank(),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig1d
ggsave("PropSexPBregression.png")
```
Including clutch size and pb*prop interaction

```{r echo=FALSE}

lm4 <- lm(Track~pb*WCSPprop+Pre+Sex+Clutch+as.numeric(Feather),data=allpbs)
summary(lm4)
lm4red <- lm(Track~pb+WCSPprop+Pre+Sex+Clutch+as.numeric(Feather),data=allpbs)
anova(lm4,lm4red)
summary(lm4red)
allpbs <- subset(allpbs,allpbs$Sex!="NA")
allpbs$figtrt <- as.factor(paste(allpbs$Sex,allpbs$pb,sep = ""))
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```
Interaction between sex and proportion (splitting up data into GC and WC trials). Since this experiment isn't really about species recognition, I think it's appropriate to focus on each species response separately from now on.

```{r echo=FALSE}

GC_sex_mod <- lm(Track~WCSPprop*Sex+as.numeric(Feather),data=gcsppbs)  
summary(GC_sex_mod)

GC_sexplus_mod <- lm(Track~WCSPprop*Sex+Clutch+Pre+as.numeric(Feather),data=gcsppbs)  
summary(GC_sexplus_mod)

fig2a <- ggplot(data=gcsppbs,aes(y=Track,x=WCSPprop,shape=Sex,fill=Sex),colour=Sex)+geom_point()+geom_smooth(method=lm, se=FALSE, fullrange=TRUE,aes(col=Sex))+scale_shape_manual(values=c(21:24))
fig2a=fig2a+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig2a+labs(y="Response to GCSP")
```

```{r echo=FALSE}
WC_sex_mod <- lm(Track~WCSPprop*Sex+as.numeric(Feather),data=wcsppbs) 
summary(WC_sex_mod)
fig2b <- ggplot(data=wcsppbs,aes(y=Track,x=WCSPprop,shape=Sex,fill=Sex),colour=Sex)+geom_point()+geom_smooth(method=lm, se=FALSE, fullrange=TRUE,aes(col=Sex))+scale_shape_manual(values=c(21:24))
fig2b=fig2b+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig2b+labs(y="Response to WCSP")
# GC_sex_clutch_mod <- lm(Track~WCSPprop*Sex+Pre+Clutch+as.numeric(Feather),data=gcsppbs)  
# summary(GC_sex_clutch_mod)
# WC_sex_clutch_mod <- lm(Track~WCSPprop*Sex+Pre+Clutch+as.numeric(Feather),data=wcsppbs) 
# summary(WC_sex_clutch_mod)
```

How about splitting into males and females?

```{r echo=FALSE}
malemod <- lm(Track~pb+Pre+WCSPprop+as.numeric(Feather), data=allpbsM)
summary(malemod)

malemod1 <- lm(Track~pb*WCSPprop+Pre+as.numeric(Feather), data=allpbsM)
summary(malemod1)

fig3a <- ggplot(data=allpbsM,aes(y=Track,x=WCSPprop,shape=pb,fill=pb),colour=pb)+geom_point()+geom_smooth(method=lm, se=FALSE, fullrange=TRUE,aes(col=pb))+scale_shape_manual(values=c(21:24))
fig3a=fig3a+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig3a+labs(y="Male responses by exposure")
ggsave("MaleResponses.png")

femalemod <- lm(Track~pb+Pre+WCSPprop+as.numeric(Feather), data=allpbsF)
summary(femalemod)
femalemod1 <- lm(Track~pb*WCSPprop+Pre+as.numeric(Feather), data=allpbsF)
summary(femalemod1)
fig3b <- ggplot(data=allpbsF,aes(y=Track,x=WCSPprop,shape=pb,fill=pb),colour=pb)+geom_point()+geom_smooth(method=lm, se=FALSE, fullrange=TRUE,aes(col=pb))+scale_shape_manual(values=c(21:24))
fig3b=fig3b+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig3b+labs(y="Female responses by exposure")
ggsave("FemaleResponses.png")
```
### Analysis of 2017 WCSP nests
```{r echo=FALSE}
wcnestdata <- read.csv(paste0(workingdir,"2017_NestlingPBs_WConly.csv"))

colnames(wcnestdata)<- c("band.number","date","time","Species","Nest.number","Feather","pb","Pre","Track","Post","Ctrl","ctrl.pre","ctrl.track","ctrl.post","notes")
wcnestdata$pb <- as.character(wcnestdata$pb)
wcnestdata[grep("WC*",wcnestdata$pb),]$pb<-"WC"
wcnestdata[grep("GC*",wcnestdata$pb),]$pb<-"GC"

fig4 =ggplot(data=wcnestdata,aes(y=Track,x=pb,fill=as.factor(Nest.number)))+geom_point(shape=21)
fig4=fig4+theme(
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
)
fig4+labs(y="WCSP nestlings")

```
